---
title: "Shell_Pest_2025_Cliona"
output: html_document
---
`r Sys.Date()`

### To make things cleaner this is only code for Cliona Intensity and Prevalence. Clams and Polydora will be addressed in separate scripts.  

```{r VariableSet, echo = FALSE, warning = FALSE, message = FALSE}
# Set your variables
Estuaries <- c("AB") # Limits to only Apalachicola Bay
Database = "OysterLocalMD20250821"  # Set the local database to use
Server = "localhost\\LOCALSQL" # Set the local Server to use
```

```{r PackageLoad, echo = FALSE, warning = FALSE, message = FALSE}
# Load necessary R packages
library(tidyverse)
library(odbc)
library(DBI)
library(dbplyr)
library(lubridate)
library(knitr)
library(openxlsx)
library(car)

#chatGPT
library(gridExtra)
library(multcompView)

# From Sarah/Colin's CI GamaAnalysis
library(Matrix)
library(glmmTMB)
library(DHARMa)
library(emmeans)
library(AICcmodavg)
library(ggplot2)
library(dplyr)
library(cowplot)
library(ggridges)
library(ggh4x)
library(DHARMa)

# From Claude
library(performance)

```

```{r ConfigureChunks, warning=FALSE, include=FALSE}
# Configure chunks
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

```{r Functions}

# Function for creating a nuce table for Bias Testing
importance_table <- function(full_model, predictors) {
  # intercept-only null model
  null_mod <- update(full_model, . ~ 1)
  
  # Full model pseudo-R²
  pseudoR2 <- function(model, null_model = NULL) {
    ll_full <- as.numeric(logLik(model))
    if (is.null(null_model)) {
      null_model <- update(model, . ~ 1)
    }
    ll_null <- as.numeric(logLik(null_model))
    1 - (ll_full / ll_null)
  }
  R2_full <- pseudoR2(full_model, null_mod)
  
  # Function for each predictor
  drop_one_safe <- function(var, full) {
    # reduced model
    reduced <- tryCatch(
      update(full, as.formula(paste(". ~ . -", var))),
      error = function(e) NULL,
      warning = function(w) {invokeRestart("muffleWarning")}
    )
    if (is.null(reduced)) {
      return(data.frame(
        Predictor = var,
        deltaAIC = NA, LRT = NA, df_diff = NA, p_value = NA,
        R2_full = R2_full, R2_reduced = NA, deltaR2 = NA,
        stringsAsFactors = FALSE
      ))
    }
    
    # logLik + AIC
    ll_full <- as.numeric(logLik(full))
    ll_red  <- as.numeric(logLik(reduced))
    aic_full <- AIC(full)
    aic_red  <- AIC(reduced)
    deltaAIC <- aic_red - aic_full
    
    # parameter df
    df_full <- attr(logLik(full), "df")
    df_red  <- attr(logLik(reduced), "df")
    if (is.null(df_full)) df_full <- length(coef(full)$cond)
    if (is.null(df_red))  df_red  <- length(coef(reduced)$cond)
    
    df_diff <- as.integer(round(df_full - df_red))
    LRT <- 2 * (ll_full - ll_red)
    p_value <- if (!is.na(df_diff) && df_diff > 0) {
      pchisq(LRT, df = df_diff, lower.tail = FALSE)
    } else {
      NA_real_
    }
    
    # pseudo R²
    R2_reduced <- pseudoR2(reduced, null_mod)
    deltaR2 <- R2_full - R2_reduced
    
    data.frame(
      Predictor = var,
      deltaAIC = deltaAIC,
      LRT = LRT,
      df_diff = df_diff,
      p_value = p_value,
      R2_full = R2_full,
      R2_reduced = R2_reduced,
      deltaR2 = deltaR2,
      stringsAsFactors = FALSE
    )
  }
  
  # Build table
  results <- do.call(rbind, lapply(predictors, drop_one_safe, full = full_model))
  results <- results %>% arrange(desc(deltaR2))
  return(results)
}

```


```{r DatabaseDownload}
# Connect to Local database server and pull all necessary data, then close connection 
con <- dbConnect(odbc(),
                    Driver = "SQL Server", 
                    Server = Server,
                    Database = Database,
                    Authentication = "ActiveDirectoryIntegrated")

hsdbShellPest <- tbl(con,in_schema("hsdb", "ShellPest")) %>%
  collect() %>%
  filter(substring(SampleEventID, 1, 2) %in% Estuaries)

hsdbConditionIndex <- tbl(con,in_schema("hsdb", "ConditionIndex")) %>%
  collect() %>%
  filter(substring(SampleEventID, 1, 2) %in% Estuaries)

DBI::dbDisconnect(con)
```

```{r DataFiltersBasic}
# Adjust data to include necessary fields and exclude incomplete years.
ShellPest <- hsdbShellPest %>% 
  mutate(Section = substring(OysterID, 3, 3),
         FixedLocation = substring(SampleEventID, 19, 22),
         Year = substring(SampleEventID, 8, 11),
         Month = substring(SampleEventID, 12, 13),
         Day = substr(SampleEventID, 14, 15),
         Date = as.Date(paste(Year, Month, Day, sep = "-")),
         Season = case_when(
          month(Date) >= 1 & month(Date) <= 3 ~ "Winter",
          month(Date) >= 4 & month(Date) <= 6 ~ "Spring",
          month(Date) >= 7 & month(Date) <= 9 ~ "Summer",
          month(Date) >= 10 & month(Date) <= 12 ~ "Fall",
          TRUE ~ "Invalid")) %>%
  left_join(hsdbConditionIndex, by = c("OysterID")) %>%
  filter(Year < 2023) %>%
  select(OysterID,
         PhotoSurface,
         TotalArea,
         ClionaArea,
         Section,
         FixedLocation,
         Date,
         Month,
         Year,
         Season,
         ProcessedBy,
         ShellHeight,
         TotalWeight) 

# Adjust to provide the sum values needed for calculations
ShellPest1 <- ShellPest %>%
  group_by(OysterID,
           Section,
           FixedLocation,
           Date,
           Month,
           Year,
           Season,
           ProcessedBy,
           ShellHeight,
           TotalWeight) %>%
  summarise(TotalSurfaceArea = sum(TotalArea, na.rm = TRUE),
         TotalExternalArea = sum(
           if_else(substring(PhotoSurface, 2, 2) == "E", TotalArea, 0), na.rm = TRUE),
         TotalInternalArea = sum(
           if_else(substring(PhotoSurface, 2, 2) == "I", TotalArea, 0), na.rm = TRUE),
         TotalTopArea = sum(
           if_else(substring(PhotoSurface, 1, 1) == "T", TotalArea, 0), na.rm = TRUE),
         TotalBottomArea = sum(
           if_else(substring(PhotoSurface, 1, 1) == "B", TotalArea, 0), na.rm = TRUE),
        # Cliona values area
         TotalClionaArea = sum(ClionaArea, na.rm = TRUE),
         ClionaExternalArea = sum(
           if_else(substring(PhotoSurface, 2, 2) == "E", ClionaArea, 0), na.rm = TRUE),
         ClionaInternalArea = sum(
           if_else(substring(PhotoSurface, 2, 2) == "I", ClionaArea, 0), na.rm = TRUE),
         ClionaTopArea = sum(
           if_else(substring(PhotoSurface, 1, 1) == "T", ClionaArea, 0), na.rm = TRUE),
         ClionaBottomArea = sum(
           if_else(substring(PhotoSurface, 1, 1) == "B", ClionaArea, 0), na.rm = TRUE),
         .groups = "drop"   # removes grouping after summarise
         ) 

ShellPest2 <- ShellPest1 %>% # getting everything that needs to be scaled
  mutate(TotalWeight_scaled = scale(TotalWeight)[,1],
         ShellHeight_scaled = scale(ShellHeight)[,1],
         Year_scaled = scale(as.numeric(Year))[,1]  # Scaled for better convergence
          ) %>% # getting everything that needs to be factored or numeric
  mutate(Section = factor(Section, levels = c("W", "C", "E")),
         Season = factor(Season, levels = c("Winter", "Spring", "Summer", "Fall")),
         Month = factor(Month, levels = 1:12, labels = month.abb),
         Year = factor(Year),   # Categorical for non-linear trends
         Year_numeric = as.numeric(as.character(Year)),  # For linear trends
         ProcessedBy = factor(ProcessedBy),
         FixedLocation = factor(FixedLocation),
         # Adjust to provide the calculated values for intensity
         ClionaIntensityWholeOyster = if_else(TotalClionaArea == 0, 0, 
                                              round(TotalClionaArea / TotalSurfaceArea, 4)),
         ClionaIntensityExternal = if_else(ClionaExternalArea == 0, 0, 
                                              round(ClionaExternalArea / TotalExternalArea, 4)),
         ClionaIntensityInternal = if_else(ClionaInternalArea == 0, 0, 
                                              round(ClionaInternalArea / TotalInternalArea, 4)),
         ClionaIntensityTop = if_else(ClionaTopArea == 0, 0, 
                                              round(ClionaTopArea / TotalTopArea, 4)),
         ClionaIntensityBottom = if_else(ClionaBottomArea == 0, 0, 
                                              round(ClionaBottomArea / TotalBottomArea, 4))
         )

# Create binary infection status (0 = uninfected, 1 = infected)
ShellPest2$Infected <- ifelse(ShellPest2$ClionaIntensityWholeOyster > 0, 1, 0)

# Remove unnecessary dataframes
rm(hsdbConditionIndex, hsdbShellPest, ShellPest, ShellPest1)

```

<br>~~~ Intensity: Initial GLMM Models ~~~ <br>

``` {r Intensity: Initial GLMM Models}

# Additional optimization options if convergence issues occur:
# control = glmmTMBControl(
#   optimizer = optim,
#   optArgs = list(method = "BFGS"),
#   start = list(theta = log(1)),  # Starting values for dispersion
#   map = list(),  # Fix parameters if needed
#   optCtrl = list(eval.max = 1000, iter.max = 1000)
# )

# Data exploration to see what distribution to use
# ggplot(ShellPest2, aes(x=ClionaIntensityWholeOyster)) + geom_histogram()
# ggplot(ShellPest2, aes(x=Infected)) + geom_histogram()

##### 1-Section ONLY as a fixed effect
model.S <- glmmTMB(ClionaIntensityWholeOyster ~ Section,
      family = beta_family(link = "logit"),
  ziformula = ~ 0,
  data = subset(ShellPest2, ClionaIntensityWholeOyster > 0))

##### 2-TotalWeight ONLY as a fixed effect 
model.TW <- glmmTMB(ClionaIntensityWholeOyster ~ TotalWeight_scaled,
       family = beta_family(link = "logit"),
  ziformula = ~ 0,
  data = subset(ShellPest2, ClionaIntensityWholeOyster > 0))

##### 3-ShellHeight ONLY as a fixed effect 
model.SH <- glmmTMB(ClionaIntensityWholeOyster ~ ShellHeight_scaled,
      family = beta_family(link = "logit"),
  ziformula = ~ 0,
  data = subset(ShellPest2, ClionaIntensityWholeOyster > 0))

##### 4-Section AND TotalWeight as a fixed effect 
model.S.TW <- glmmTMB(ClionaIntensityWholeOyster ~ Section + TotalWeight_scaled,
    family = beta_family(link = "logit"),
  ziformula = ~ 0,
  data = subset(ShellPest2, ClionaIntensityWholeOyster > 0))

##### 5-Section AND ShellHeight as a fixed effect 
model.S.SH <- glmmTMB(ClionaIntensityWholeOyster ~ Section + ShellHeight_scaled,
    family = beta_family(link = "logit"),
  ziformula = ~ 0,
  data = subset(ShellPest2, ClionaIntensityWholeOyster > 0))

##### 6-TotalWeight AND ShellHeight as a fixed effect 
model.TW.SH <- glmmTMB(ClionaIntensityWholeOyster ~ TotalWeight_scaled + ShellHeight_scaled,
    family = beta_family(link = "logit"),
  ziformula = ~ 0,
  data = subset(ShellPest2, ClionaIntensityWholeOyster > 0))

##### 7-TotalWeight AND ShellHeight AND Section as a fixed effect 
model.TW.SH.S <- glmmTMB(
  ClionaIntensityWholeOyster ~ TotalWeight_scaled + ShellHeight_scaled + Section,
     family = beta_family(link = "logit"),
  ziformula = ~ 0,
  data = subset(ShellPest2, ClionaIntensityWholeOyster > 0))

### Use AICc to compare the 2 models: we'll use the model with the lowest score for making any inferences moving forward
Intensity_models1 <- list(
  '1-SectionOnly' = model.S,
  '2-TotalWeightOnly' = model.TW, 
  '3-ShellHeightOnly' = model.SH, 
  '4-Section+TotalWeight' = model.S.TW, 
  '5-Section+SH' = model.S.SH,
  '6-TotalWeight+SH' = model.TW.SH,
  '7-TotalWeight+SH+Section' = model.TW.SH.S)

aictab(Intensity_models1) #will use one for analysis

# Key findings from model selection
cat("=== Intensity: INITIAL MODEL SELECTION INSIGHTS === 
      1. Section is highly important (1 >> 2 or 3 alone), TotalWeight is more important than ShellHeight (2 >> 3).
      2. The best model includes all three effects (7 >> 4 or 5 or 6).
      3. The best model fit involved subsetting the data to exclude oysters with no Cliona. Including cliona-free oysters caused issues with residuals due to the large number of zeros in the data.")

######### Since 7-TotalWeight+SH+Section has a Massively better AICc, only it will be tested here
best_model1 <- model.TW.SH.S
#########

cat('The following is Residual results, QQ Plot, R2, and Summary for the best fitting model ONLY: 
    7-TotalWeight+SH+Section')

##### Assess goodness-of-fit
simulateResiduals(best_model1, n = 250, refit = FALSE, plot = TRUE)

# Calculate R-squared equivalent for Intensity models. Closer to 1 is better
r2(best_model1)

##### Look at estimates
summary(best_model1)

# Key findings from Initial model 
cat("=== Intensity: INITIAL MODEL INSIGHTS ===
  ClionaIntensityWholeOyster ~ TotalWeight_scaled + ShellHeight_scaled + Section
      1. Model fits the data alright with decent residuals. KS and disperison tests indicate fit issues, though this is common with large data sets.
      2. R-squared of 0.104 indicates a weak predictive ability, meaning there's lots of room for improvement.
      2. Larger oysters have more a higher intensity of cliona This makes sense and conforms to basic observations.
      4. Counterintuitively, oysters with a greater shell height show slightly less cliona at a given weight.
      3. There are statistically significant differences in the cliona intensity observed in the sections of the Bay.")

# Remove unnecesssary models
rm(model.S, model.S.SH, model.S.TW, model.SH, model.TW, model.TW.SH, model.TW.SH.S)

```

<br>~~~ Intensity: Improving Basic GLMM Models ~~~ <br>
Adding Season and Year to best Initial model

``` {r Intensity: Improving Basic GLMM Models}
# Extended model set: Adding Season and Year systematically

##### 1-Baseline - Best model from above
### TotalWeight AND ShellHeight AND Section as a fixed effect 
# model.TW.SH.S <- glmmTMB(
#   ClionaIntensityWholeOyster ~ TotalWeight_scaled + ShellHeight_scaled + Section,
#      family = beta_family(link = "logit"),
#   ziformula = ~ 0,
#   data = subset(ShellPest2, ClionaIntensityWholeOyster > 0))

# 2-Add Season (cyclical pattern)
# Section AND TotalWeight AND Season as a fixed effect 
model.TW.SH.S.Ss <- glmmTMB(ClionaIntensityWholeOyster ~ TotalWeight_scaled + ShellHeight_scaled + Section + Season,
                   family = beta_family(link = "logit"),
  ziformula = ~ 0,
  data = subset(ShellPest2, ClionaIntensityWholeOyster > 0))

# 3-Add Year as a factor (non-linear temporal trend)
# Section AND TotalWeight AND Year as a fixed effect 
model.TW.SH.S.Y <- glmmTMB(ClionaIntensityWholeOyster ~ TotalWeight_scaled + ShellHeight_scaled + Section + Year,
                   family = beta_family(link = "logit"),
  ziformula = ~ 0,
  data = subset(ShellPest2, ClionaIntensityWholeOyster > 0))

# 4-Add Year as a linear trend (more parsimonious)
# Section AND TotalWeight AND Year_scaled as a fixed effect 
model.TW.SH.S.Y_s <- glmmTMB(ClionaIntensityWholeOyster ~ TotalWeight_scaled + ShellHeight_scaled + Section + Year_scaled,
                   family = beta_family(link = "logit"),
  ziformula = ~ 0,
  data = subset(ShellPest2, ClionaIntensityWholeOyster > 0))

# 5-Add Season AND Year (factor)
# Section AND TotalWeight AND Season AND Year as a fixed effect 
model.TW.SH.S.Ss.Y <- glmmTMB(ClionaIntensityWholeOyster ~ TotalWeight_scaled + ShellHeight_scaled + Section + Season + Year,
                   family = beta_family(link = "logit"),
  ziformula = ~ 0,
  data = subset(ShellPest2, ClionaIntensityWholeOyster > 0))

# 6-Add Season AND Year (linear)
# Section AND TotalWeight AND Season AND Year_Scaled as a fixed effect 
model.TW.SH.S.Ss.Y_s <- glmmTMB(ClionaIntensityWholeOyster ~ TotalWeight_scaled + ShellHeight_scaled + Section + Season + Year_scaled,
                   family = beta_family(link = "logit"),
  ziformula = ~ 0,
  data = subset(ShellPest2, ClionaIntensityWholeOyster > 0))

# 7-Add Season-Year interaction (test if seasonal patterns change over years)
# Section AND TotalWeight AND Season-Year_scaled interaction as a fixed effect 
model.TW.SH.S.Ss.X.Y_s <- glmmTMB(ClionaIntensityWholeOyster ~ TotalWeight_scaled + ShellHeight_scaled + Section + Season * Year_scaled,
                   family = beta_family(link = "logit"),
  ziformula = ~ 0,
  data = subset(ShellPest2, ClionaIntensityWholeOyster > 0))

### Use AICc to compare the 2 models: we'll use the model with the lowest score for making any inferences moving forward
Intensity_models2 <- list('1-Section+TotalWeight+ShellHeight (Base)' = best_model1, 
                 '2-SEC+TW+SH+Season' = model.TW.SH.S.Ss, 
                 '3-SEC+TW+SH+Year' = model.TW.SH.S.Y, 
                 '4-SEC+TW+SH+YearScaled' = model.TW.SH.S.Y_s, 
                 '5-SEC+TW+SH+Season+Year' = model.TW.SH.S.Ss.Y,
                 '6-SEC+TW+SH+Season+YearScaled' = model.TW.SH.S.Ss.Y_s,
                 '7-SEC+TW+SH+Season*YearScaled' = model.TW.SH.S.Ss.X.Y_s
)

aictab(Intensity_models2) #will use one for analysis


# Key findings from model selection
cat("=== Counts: IMPROVED MODEL SELECTION INSIGHTS ===
      1. Both Season AND Year are highly important (5 >> 2 or 3 alone)
      2. Non-linear year effects are crucial (5 >> 6: factor vs linear year)
      3. Seasonal patterns are consistent across years (5 >> 7: no interaction needed)
      4. Temporal factors explain MASSIVE variation (ΔAICc = 141 vs baseline)")

######### Since 5-SEC+TW+SH+Season+Year has a significantly better AICc, only it will be tested here
best_model2 <- model.TW.SH.S.Ss.Y
#########

cat('Following is Residual results, QQ Plot, ACF series residuals, R2, and Summary for the best fitting model ONLY:   
    5-SEC+WT+Season+Year')

##### Assess goodness-of-fit
simulateResiduals(best_model2, n = 175, refit = FALSE, plot = TRUE)

# Check for temporal autocorrelation since we have time series data
acf(residuals(best_model2, type = "pearson"))
# The residuals exhibit no significant autocorrelation, implying that the model has effectively captured the temporal dependencies. 

# Calculate R-squared equivalent for count models. Closer to 1 is better
r2(best_model2)

##### Look at estimates
summary(best_model2)

# Key findings from Improved model 
cat("=== Intensity: IMPROVED MODEL INSIGHTS ===
  ClionaIntensityWholeOyster ~ TotalWeight_scaled + ShellHeight_scaled + Section + Season + Year
      1. Model fits the data pretty well with good residuals and and an improved R-squared (0.155) than the Initial model. Thus there is still a good amount of unexpalined variability.
      2. ACF series residuals series exhibit no significant autocorrelation, implying that the model has effectively captured the temporal dependencies.
      3. Key insights about larger oysters and differences between sections still hold.
      4. There appear to be some statistical differences between the Seasons and the Years. Closer inspection of those results to follow.")

# Remove unnecessary models

rm(model.TW.SH.S.Ss, model.TW.SH.S.Y, model.TW.SH.S.Y_s, model.TW.SH.S.Ss.Y, model.TW.SH.S.Ss.Y_s, model.TW.SH.S.Ss.X.Y_s)

```

<br>~~~ Intensity: Bias Testing ~~~ <br>
Examining the impact of observers and locations, as both have changed over the years

``` {r Intensity: Bias Testing}

# BIAS TESTING MODELS

##### 1-Baseline - Best model from above
### Add Season AND Year (factor)
# # Section AND TotalWeight AND Season AND Year as a fixed effect 
# model.TW.SH.S.Ss.Y <- glmmTMB(ClionaIntensityWholeOyster ~ TotalWeight_scaled + ShellHeight_scaled + Section + Season + Year,
#                    family = beta_family(link = "logit"),
#   ziformula = ~ 0,
#   data = subset(ShellPest2, ClionaIntensityWholeOyster > 0))

# OBSERVER BIAS TESTING 
# 2-Add ProcessedBy to Intensity model
model.Intensity_observer <- glmmTMB(ClionaIntensityWholeOyster ~ TotalWeight_scaled + ShellHeight_scaled + Section + Season + Year + ProcessedBy,
                            family = beta_family(link = "logit"),
  ziformula = ~ 0,
  data = subset(ShellPest2, ClionaIntensityWholeOyster > 0))

# LOCATION BIAS TESTING
# 3-Add FixedLocation as random effect (nested within Section) to Intensity model
model.Intensity_location <- glmmTMB(ClionaIntensityWholeOyster ~ TotalWeight_scaled + ShellHeight_scaled + Section + Season + Year + (1|FixedLocation),
                            family = beta_family(link = "logit"),
  ziformula = ~ 0,
  data = subset(ShellPest2, ClionaIntensityWholeOyster > 0))

# COMBINED BIAS TESTING 
# 4-Both observer and location effects on Intensity
model.Intensity_both <- glmmTMB(ClionaIntensityWholeOyster ~ TotalWeight_scaled + ShellHeight_scaled + Section + Season + Year + ProcessedBy + (1|FixedLocation),
                        family = beta_family(link = "logit"),
  ziformula = ~ 0,
  data = subset(ShellPest2, ClionaIntensityWholeOyster > 0))

### Use AICc to compare the 2 models: we'll use the model with the lowest score for making any inferences moving forward
# For Intensity models
Intensity_bias_models <- list(
  "Original" = best_model2,  # Your best Improved model
  "Observer_Bias" = model.Intensity_observer,
  "Location_Bias" = model.Intensity_location, 
  "Both_Biases" = model.Intensity_both
)

aictab(Intensity_bias_models)

# Key findings from model selection
cat("=== Intensity: BIAS TESTING MODEL SELECTION INSIGHTS ===
      1. Both Biases are present in Cliona Intensity data (ΔAICc = 114 vs baseline)
      2. Location and Observer have a MASSIVE impact on results")

# simulateResiduals(model.Intensity_observer, n = 150, refit = FALSE, plot = TRUE)
# simulateResiduals(model.Intensity_location, n = 150, refit = FALSE, plot = TRUE)
simulateResiduals(model.Intensity_both, n = 150, refit = FALSE, plot = TRUE)

# Check for temporal autocorrelation since we have time series data
acf(residuals(model.Intensity_both, type = "pearson"))

# r2(model.Intensity_observer)
# r2(model.Intensity_location)
r2(model.Intensity_both)

summary(model.Intensity_both)

# Key findings from Bias testing model 
cat("=== Intensity: BIAS TESTING MODEL INSIGHTS ===
  ClionaIntensityWholeOyster ~ TotalWeight_scaled + ShellHeight_scaled + Section + Season + Year + ProcessedBy + (1|FixedLocation)
      1. Model fits the data well with good residuals and an R-squared much higher than the Improved model. KS Test and Dispersion fail as expected with a large data set, and nothing looks wrong with the plot otherwise.
      2. ACF series residuals series exhibit no significant autocorrelation, implying that the model has effectively captured the temporal dependencies.
      3. The MASSIVE increase in model performance with both biases raise a lot of doubt about the Improved Model's insights. ")

# Assign predictors of variance to see what has the most amount of influence in the bias model
predictors <- c("Section", "TotalWeight_scaled", "ShellHeight_scaled", "Season", "Year", "ProcessedBy")

# Uses Intraclass Correlation Coefficient (ICC) to quantify how much variation in your outcomeis attributable to Random effects
cat("Intraclass Correlation Coefficient (ICC) quantifies how much variation is attributable to Random effects:")
icc(model.Intensity_both)

# uses the importantance_table function to make a clean table detail each factor's importance in the final model
importance_results <- importance_table(model.Intensity_both, predictors) 
kable(importance_results, format = "pipe", digits = 3)

# Key findings from Bias Importance  
cat("=== Intensity: BIAS TESTING IMPORTANCE TABLE & ICC ===
      1. Adjusted ICC indicates that after controliing for other factors, ~15% of the variance is attributable to the random effect of FixedLocation. Thus it is not the main driver of variation but it is meaningful.
      2. Large Positive ΔAIC indicates that TotalWeight and Processed by are extremely important predictors. 
      3. Section's negative ΔAIC indicates it doesn't really improve the model.
      4. TotalWeight and Processed by have very small p-values (e-16 and e-7), indicating they are highly significant.
      5. Season and Year are both significant with very low p-values (e-6 and 0.002), but far less than weight and processor.
      6. Shell Height effects are weak, but statistically significant (p = 0.04)
      7. Section is not significant (p = 0.16).
      8. ΔR2 measures how much of the model’s explained variance is uniquely attributable to that predictor. Even the most important variables explain a tiny amount of variance (less than 2%).")

cat(" |||||||||||||||||||||||||||| CONCLUSION |||||||||||||||||||||||||||| 
    1. TotalWeight is by far the strongest driver of cliona intensity. Heavier oysters tend to have more clams.
    2. Given the outsized impact of bias on this data, no other strong insights can be drawn without significant reprocessing, retraining, and robust QA measures being put in place. After which all these tests need to be run again and more detailed analysis run to determine any significant differences.
    3. This solution represents a large input of staff time (thousands of hours) for a small amount of benefit.
*** 4. Recommendation is to discontinue measuring cliona intensity in ShellPest data. ***")

```

<br>~~~ Prevalence: Initial GLMM Models ~~~ <br>

``` {r Prevalence: Initial GLMM Models}

# Additional optimization options if convergence issues occur:
# control = glmmTMBControl(
#   optimizer = optim,
#   optArgs = list(method = "BFGS"),
#   start = list(theta = log(1)),  # Starting values for dispersion
#   map = list(),  # Fix parameters if needed
#   optCtrl = list(eval.max = 1000, iter.max = 1000)
# )

##### 1-Section ONLY as a fixed effect
P.model.S <- glmmTMB(Infected ~ Section,
                   family = binomial,
                   data = ShellPest2,
                   control = glmmTMBControl(optimizer = optim, 
                                          optArgs = list(method = "BFGS")))

##### 2-TotalWeight ONLY as a fixed effect 
P.model.TW <- glmmTMB(Infected ~ TotalWeight_scaled,
                   family = binomial,
                   data = ShellPest2,
                   control = glmmTMBControl(optimizer = optim, 
                                          optArgs = list(method = "BFGS")))

##### 3-Section AND ShellHeight as a fixed effect 
P.model.SH <- glmmTMB(Infected ~ ShellHeight_scaled,
                   family = binomial,
                   data = ShellPest2,
                   control = glmmTMBControl(optimizer = optim, 
                                          optArgs = list(method = "BFGS")))

##### 4-Section AND TotalWeight as a fixed effect 
P.model.S.TW <- glmmTMB(Infected ~ Section + TotalWeight_scaled,
                   family = binomial,
                   data = ShellPest2,
                   control = glmmTMBControl(optimizer = optim, 
                                          optArgs = list(method = "BFGS")))

##### 5-Section AND ShellHeight as a fixed effect 
P.model.S.SH <- glmmTMB(Infected ~ Section + ShellHeight_scaled,
                   family = binomial,
                   data = ShellPest2,
                   control = glmmTMBControl(optimizer = optim, 
                                          optArgs = list(method = "BFGS")))

##### 6-TotalWeight AND ShellHeight as a fixed effect 
P.model.TW.SH <- glmmTMB(Infected ~ TotalWeight_scaled + ShellHeight_scaled,
                   family = binomial,
                   data = ShellPest2,
                   control = glmmTMBControl(optimizer = optim, 
                                          optArgs = list(method = "BFGS")))

##### 7-TotalWeight AND ShellHeight AND Section as a fixed effect 
P.model.TW.SH.S <- glmmTMB(Infected ~ TotalWeight_scaled + ShellHeight_scaled + Section,
                   family = binomial,
                   data = ShellPest2,
                   control = glmmTMBControl(optimizer = optim, 
                                          optArgs = list(method = "BFGS")))

### Use AICc to compare the 2 models: we'll use the model with the lowest score for making any inferences moving forward
Prevalence_models1 <- list('1-SectionOnly' = P.model.S, 
                 '2-TotalWeightOnly' = P.model.TW, 
                 '3-ShellHeightOnly' = P.model.SH, 
                 '4-Section+TotalWeight' = P.model.S.TW, 
                 '5-Section+SH' = P.model.S.SH,
                 '6-TotalWeight+SH' = P.model.TW.SH,
                 '7-TotalWeight+SH+Section' = P.model.TW.SH.S)

aictab(Prevalence_models1) #will use one for analysis

# Key findings from model selection
cat("=== Prevalence: INITIAL MODEL SELECTION INSIGHTS === 
      1. Adding ShellHeight to Model #4 improved AICc slightly, with worse residuals, and no change to R2 value. Thus, preference for the simpler model means that #4 was chosen as the best model.
      2. Section is highly important (1 >> 2 or 3)
      3. Of the two, TotalWeight is more important than ShellHeight (2 > 3)")

######### Since Section+TotalWeight has a Massively better AICc, only it will be tested here
P.best_model1 <- P.model.S.TW
#########

##### Assess goodness-of-fit
simulateResiduals(P.best_model1, n = 250, refit = FALSE, plot = TRUE)

# Calculate R-squared equivalent for count models. Closer to 1 is better
r2(P.best_model1)

##### Look at estimates
summary(P.best_model1)

# Key findings from Initial model 
cat("=== Prevalence: INITIAL MODEL INSIGHTS ===
  Infected ~ Section + TotalWeight_scaled
      1. Model fits the data pretty well with good residuals.
      2. Tjur's R-squared is used here because Prevalence is a binary outcome. A value of 0.294 indicates a moderate ability to discriminate between the 2 outcomes. Thus, there's room for improvement.
      3. Larger oysters have a higher prevalence. This makes sense and conforms to basic observations.
      4. There are statistically significant differences in the prevalence of cliona infestation observed in the sections of the Bay.")

# Remove unnecessay models
rm(P.model.S, P.model.TW, P.model.SH, P.model.S.TW, P.model.S.SH, P.model.TW.SH, P.model.TW.SH.S)

```


<br>~~~ Prevalence: Improving Basic GLMM Models ~~~ <br>
Adding Season and Year to best Initial model

``` {r Prevalence: Improving Basic GLMM Models}
# Extended model set: Adding Season and Year systematically

##### 1-Baseline - Best model from above
## Section AND TotalWeight as a fixed effect 
# P.model.S.TW <- glmmTMB(Infected ~ Section + TotalWeight_scaled,
#                    family = binomial,
#                    data = ShellPest2,
#                    control = glmmTMBControl(optimizer = optim, 
#                                           optArgs = list(method = "BFGS")))

# 2-Add Season (cyclical pattern)
# Section AND TotalWeight AND Season as a fixed effect 
P.model.S.TW.Ss <- glmmTMB(Infected ~ Section + TotalWeight_scaled + Season,
                   family = binomial,
                   data = ShellPest2,
                   control = glmmTMBControl(optimizer = optim, 
                                          optArgs = list(method = "BFGS")))

# 3-Add Year as a factor (non-linear temporal trend)
# Section AND TotalWeight AND Year as a fixed effect 
P.model.S.TW.Y <- glmmTMB(Infected ~ Section + TotalWeight_scaled + Year,
                   family = binomial,
                   data = ShellPest2,
                   control = glmmTMBControl(optimizer = optim, 
                                          optArgs = list(method = "BFGS")))

# 4-Add Year as a linear trend (more parsimonious)
# Section AND TotalWeight AND Year_scaled as a fixed effect 
P.model.S.TW.Y_s <- glmmTMB(Infected ~ Section + TotalWeight_scaled + Year_scaled,
                   family = binomial,
                   data = ShellPest2,
                   control = glmmTMBControl(optimizer = optim, 
                                          optArgs = list(method = "BFGS")))

# 5-Add Season AND Year (factor)
# Section AND TotalWeight AND Season AND Year as a fixed effect 
P.model.S.TW.Ss.Y <- glmmTMB(Infected ~ Section + TotalWeight_scaled + Season + Year,
                   family = binomial,
                   data = ShellPest2,
                   control = glmmTMBControl(optimizer = optim, 
                                          optArgs = list(method = "BFGS")))

# 6-Add Season AND Year (linear)
# Section AND TotalWeight AND Season AND Year_Scaled as a fixed effect 
P.model.S.TW.Ss.Y_s <- glmmTMB(Infected ~ Section + TotalWeight_scaled + Season + Year_scaled,
                   family = binomial,
                   data = ShellPest2,
                   control = glmmTMBControl(optimizer = optim, 
                                          optArgs = list(method = "BFGS")))


# 7-Add Season-Year interaction (test if seasonal patterns change over years)
# Section AND TotalWeight AND Season-Year_scaled interaction as a fixed effect 
P.model.S.TW.Ss.X.Y_s <- glmmTMB(Infected ~ Section + TotalWeight_scaled + Season * Year_scaled,
                   family = binomial,
                   data = ShellPest2,
                   control = glmmTMBControl(optimizer = optim, 
                                          optArgs = list(method = "BFGS")))

### Use AICc to compare the 2 models: we'll use the model with the lowest score for making any inferences moving forward
Prevalence_models2 <- list('1-SectionWeight (Base)' = P.best_model1, 
                 '2-SectionWeight+Season' = P.model.S.TW.Ss, 
                 '3-SectionWeight+Year' = P.model.S.TW.Y, 
                 '4-SectionWeight+YearScaled' = P.model.S.TW.Y_s, 
                 '5-SectionWeight+Season+Year' = P.model.S.TW.Ss.Y,
                 '6-SectionWeight+Season+YearScaled' = P.model.S.TW.Ss.Y_s,
                 '7-SectionWeight+Season * YearScaled' = P.model.S.TW.Ss.X.Y_s)

aictab(Prevalence_models2) #will use one for analysis

# Key findings from model selection
cat("=== Prevalence: MODEL SELECTION INSIGHTS ===
      1. Year is highly important (3 >> 5 or 2 alone)
      2. Non-linear year effects are crucial (5 >> 6: factor vs linear year)
      3. Seasonal patterns are consistent across years (5 >> 7: no interaction needed)
      4. Temporal factors explain MASSIVE variation (ΔAICc = 216 vs baseline)")

######### Since SSectionWeight+Year has a significantly better AICc, only it will be tested here
P.best_model2 <- P.model.S.TW.Y
#########


##### Assess goodness-of-fit
simulateResiduals(P.best_model2, n = 175, refit = FALSE, plot = TRUE)

# Check for temporal autocorrelation since we have time series data
acf(residuals(P.best_model2, type = "pearson"))
# The residuals exhibit no significant autocorrelation, implying that the model has effectively captured the temporal dependencies. 

# Calculate R-squared equivalent for count models. Closer to 1 is better
r2(P.best_model2)

##### Look at estimates
summary(P.best_model2)

# Key findings from Improved model 
cat("=== Prevalence: IMPROVED MODEL INSIGHTS ===
  Infected ~ Section + TotalWeight_scaled + Year
      1. Model fits the data pretty well with good residuals and and R-squared even closer to 1 than the Initial model.
      2. ACF series residuals series exhibit SIGNIFICANT autocorrelation, implying that the model has NOT captured the temporal dependencies. This is a major issue to resolve before any other insights can be drawn.")

# Remove unnecessary models
rm(P.model.S.TW.Ss, P.model.S.TW.Y, P.model.S.TW.Y_s, P.model.S.TW.Ss.Y, P.model.S.TW.Ss.Y_s, P.model.S.TW.Ss.X.Y_s)

```


<br>~~~ Prevalence: Bias Testing ~~~ <br>
Examining the impact of observers and locations, as both have changed over the years

``` {r Prevalence: Bias Testing}

# BIAS TESTING MODELS

##### 1-Baseline - Best model from above
### Add Year as a factor (non-linear temporal trend)
# # Section AND TotalWeight AND Year as a fixed effect 
# P.model.S.TW.Y <- glmmTMB(Infected ~ Section + TotalWeight_scaled + Year,
#                    family = binomial,
#                    data = ShellPest2,
#                    control = glmmTMBControl(optimizer = optim, 
#                                           optArgs = list(method = "BFGS")))




# OBSERVER BIAS TESTING 
# 2-Add ProcessedBy to prevalence model
model.P_observer <- glmmTMB(Infected ~ Section + TotalWeight_scaled + Year + ProcessedBy,
                           family = binomial,
                           data = ShellPest2,
                           control = glmmTMBControl(optimizer = optim,
                                                  optArgs = list(method = "BFGS")))


# LOCATION BIAS TESTING
# 3-Add FixedLocation as random effect (nested within Section) to prevalence model
model.P_location <- glmmTMB(Infected ~ Section + TotalWeight_scaled + Year + (1|FixedLocation),
                           family = binomial,
                           data = ShellPest2,
                           control = glmmTMBControl(optimizer = optim,
                                                  optArgs = list(method = "BFGS")))


# COMBINED BIAS TESTING 
# 4-Both observer and location effects on prevalence
model.P_both <- glmmTMB(Infected ~ Section + TotalWeight_scaled + Year + ProcessedBy + (1|FixedLocation),
                       family = binomial,
                       data = ShellPest2,
                       control = glmmTMBControl(optimizer = optim,
                                              optArgs = list(method = "BFGS")))

### Use AICc to compare the 2 models: we'll use the model with the lowest score for making any inferences moving forward
# For COUNT models
prevalence_bias_models <- list(
  "Original" = P.best_model2,  # Your best Improved model
  "Observer_Bias" = model.P_observer,
  "Location_Bias" = model.P_location, 
  "Both_Biases" = model.P_both
)

aictab(prevalence_bias_models)

# Key findings from model selection
cat("=== Prevalence: BIAS TESTING MODEL SELECTION INSIGHTS ===
      1. Both Biases are present in Cliona prevalence data (ΔAICc = 905 vs baseline)
      2. Location and Observer have a large impact on results")

# simulateResiduals(model.P_observer, n = 250, refit = FALSE, plot = TRUE)
# simulateResiduals(model.P_location, n = 250, refit = FALSE, plot = TRUE)
simulateResiduals(model.P_both, n = 250, refit = FALSE, plot = TRUE)

# Check for temporal autocorrelation since we have time series data
acf(residuals(model.P_both, type = "pearson"))

r2(model.P_both)

summary(model.P_both)

# Key findings from Bias testing model 
cat("=== Prevalence: BIAS TESTING MODEL INSIGHTS ===
  Infected ~ Section + TotalWeight_scaled + Year + ProcessedBy + (1|FixedLocation)
      1. Model fits the data pretty well with good residuals and and R-squared MASSIVELY better than the Improved model. KS Test fail is expected with a large data set, and nothing looks wrong with the plot otherwise.
      2. ACF series residuals series looks much better, exhibiting no significant autocorrelation, implying that the model has effectively captured the temporal dependencies.
      3. The MASSIVE increase in model performance with both biases raise a lot of doubt about the Improved Model's insights. ")

# Assign predictors of variance to see what has the most amount of influence in the bias model
predictors.P <- c("Section", "TotalWeight_scaled", "Year", "ProcessedBy")

# Uses Intraclass Correlation Coefficient (ICC) to quantify how much variation in your outcomeis attributable to Random effects
cat("Intraclass Correlation Coefficient (ICC) quantifies how much variation is attributable to Random effects:")
icc(model.P_both)

# uses the importantance_table function to make a clean table detail each factor's importance in the final model
importance_results.P <- importance_table(model.P_both, predictors.P) 
kable(importance_results.P, format = "pipe", digits = 3)

# Key findings from Bias Importance  
cat("=== Prevalence: BIAS TESTING IMPORTANCE TABLE & ICC ===
      1. Adjusted ICC indicates that after controliing for other factors, ~35% of the variance is attributable to the random effect of FixedLocation, which is substantial.
      2. Large Positive ΔAIC indicates that TotalWeight and Processed by are extremely important predictors. 
      3. Section and Year's low ΔAIC indicates they doesn't improve the model much.
      4. TotalWeight and Processed by have very small p-values (e-55 and e-10), indicating they are highly significant.
      5. Year is significant (p = 0.005).
      6. Section is not significant (p = 0.53).
      7. ΔR2 measures how much of the model’s explained variance is uniquely attributable to that predictor. TotalWeight explains ~4% on its own; ProcessedBy adds ~1.2%; Year and Season add less than 1%.")

cat(" |||||||||||||||||||||||||||| CONCLUSION |||||||||||||||||||||||||||| 
    1. TotalWeight is by far the strongest driver of cliona prevalence. Heavier oysters tend to have higher cliona prevalence.
    2. Given the outsized impact of bias on this data, no other strong insights can be drawn without significant reprocessing, retraining, and robust QA measures being put in place. After which all these tests need to be run again and more detailed analysis run to determine any significant differences.
    3. This solution represents a large input of staff time (thousands of hours) for a small amount of benefit.
*** 4. Recommendation is to discontinue noting the presence or absence of cliona in ShellPest data. ***")

```

