---
title: "Shell_Pest_2025_Polydora"
output: html_document
---
`r Sys.Date()`

### To make things cleaner this is only code for Polydora Intensity and Prevalence. Clams and Cliona will be addressed in separate scripts.  

```{r VariableSet, echo = FALSE, warning = FALSE, message = FALSE}
# Set your variables
Estuaries <- c("AB") # Limits to only Apalachicola Bay
Database = "OysterLocalMD20250821"  # Set the local database to use
Server = "localhost\\LOCALSQL" # Set the local Server to use
```

```{r PackageLoad, echo = FALSE, warning = FALSE, message = FALSE}
# Load necessary R packages
library(tidyverse)
library(odbc)
library(DBI)
library(dbplyr)
library(lubridate)
library(knitr)
library(openxlsx)
library(car)

#chatGPT
library(gridExtra)
library(multcompView)

# From Sarah/Colin's CI GamaAnalysis
library(Matrix)
library(glmmTMB)
library(DHARMa)
library(emmeans)
library(AICcmodavg)
library(ggplot2)
library(dplyr)
library(cowplot)
library(ggridges)
library(ggh4x)
library(DHARMa)

# From Claude
library(performance)

```

```{r ConfigureChunks, warning=FALSE, include=FALSE}
# Configure chunks
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```


```{r Functions}

# Function for creating a nuce table for Bias Testing
importance_table <- function(full_model, predictors) {
  # intercept-only null model
  null_mod <- update(full_model, . ~ 1)
  
  # Full model pseudo-R²
  pseudoR2 <- function(model, null_model = NULL) {
    ll_full <- as.numeric(logLik(model))
    if (is.null(null_model)) {
      null_model <- update(model, . ~ 1)
    }
    ll_null <- as.numeric(logLik(null_model))
    1 - (ll_full / ll_null)
  }
  R2_full <- pseudoR2(full_model, null_mod)
  
  # Function for each predictor
  drop_one_safe <- function(var, full) {
    # reduced model
    reduced <- tryCatch(
      update(full, as.formula(paste(". ~ . -", var))),
      error = function(e) NULL,
      warning = function(w) {invokeRestart("muffleWarning")}
    )
    if (is.null(reduced)) {
      return(data.frame(
        Predictor = var,
        deltaAIC = NA, LRT = NA, df_diff = NA, p_value = NA,
        R2_full = R2_full, R2_reduced = NA, deltaR2 = NA,
        stringsAsFactors = FALSE
      ))
    }
    
    # logLik + AIC
    ll_full <- as.numeric(logLik(full))
    ll_red  <- as.numeric(logLik(reduced))
    aic_full <- AIC(full)
    aic_red  <- AIC(reduced)
    deltaAIC <- aic_red - aic_full
    
    # parameter df
    df_full <- attr(logLik(full), "df")
    df_red  <- attr(logLik(reduced), "df")
    if (is.null(df_full)) df_full <- length(coef(full)$cond)
    if (is.null(df_red))  df_red  <- length(coef(reduced)$cond)
    
    df_diff <- as.integer(round(df_full - df_red))
    LRT <- 2 * (ll_full - ll_red)
    p_value <- if (!is.na(df_diff) && df_diff > 0) {
      pchisq(LRT, df = df_diff, lower.tail = FALSE)
    } else {
      NA_real_
    }
    
    # pseudo R²
    R2_reduced <- pseudoR2(reduced, null_mod)
    deltaR2 <- R2_full - R2_reduced
    
    data.frame(
      Predictor = var,
      deltaAIC = deltaAIC,
      LRT = LRT,
      df_diff = df_diff,
      p_value = p_value,
      R2_full = R2_full,
      R2_reduced = R2_reduced,
      deltaR2 = deltaR2,
      stringsAsFactors = FALSE
    )
  }
  
  # Build table
  results <- do.call(rbind, lapply(predictors, drop_one_safe, full = full_model))
  results <- results %>% arrange(desc(deltaR2))
  return(results)
}

```

```{r DatabaseDownload}
# Connect to Local database server and pull all necessary data, then close connection 
con <- dbConnect(odbc(),
                    Driver = "SQL Server", 
                    Server = Server,
                    Database = Database,
                    Authentication = "ActiveDirectoryIntegrated")

hsdbShellPest <- tbl(con,in_schema("hsdb", "ShellPest")) %>%
  collect() %>%
  filter(substring(SampleEventID, 1, 2) %in% Estuaries)

hsdbConditionIndex <- tbl(con,in_schema("hsdb", "ConditionIndex")) %>%
  collect() %>%
  filter(substring(SampleEventID, 1, 2) %in% Estuaries)

DBI::dbDisconnect(con)
```

```{r DataFiltersBasic}
# Adjust data to include necessary fields and exclude incomplete years.
ShellPest <- hsdbShellPest %>% 
  mutate(Section = substring(OysterID, 3, 3),
         FixedLocation = substring(SampleEventID, 19, 22),
         Year = substring(SampleEventID, 8, 11),
         Month = substring(SampleEventID, 12, 13),
         Day = substr(SampleEventID, 14, 15),
         Date = as.Date(paste(Year, Month, Day, sep = "-")),
         Season = case_when(
          month(Date) >= 1 & month(Date) <= 3 ~ "Winter",
          month(Date) >= 4 & month(Date) <= 6 ~ "Spring",
          month(Date) >= 7 & month(Date) <= 9 ~ "Summer",
          month(Date) >= 10 & month(Date) <= 12 ~ "Fall",
          TRUE ~ "Invalid")) %>%
  left_join(hsdbConditionIndex, by = c("OysterID")) %>%
  filter(Year < 2023) %>%
  select(OysterID,
         PhotoSurface,
         TotalArea,
         PolydoraArea,
         Section,
         FixedLocation,
         Date,
         Month,
         Year,
         Season,
         ProcessedBy,
         ShellHeight,
         TotalWeight) 

# Adjust to provide the sum values needed for calculations
ShellPest1 <- ShellPest %>%
  group_by(OysterID,
           Section,
           FixedLocation,
           Date,
           Month,
           Year,
           Season,
           ProcessedBy,
           ShellHeight,
           TotalWeight) %>%
  summarise(TotalSurfaceArea = sum(TotalArea, na.rm = TRUE),
         TotalExternalArea = sum(
           if_else(substring(PhotoSurface, 2, 2) == "E", TotalArea, 0), na.rm = TRUE),
         TotalInternalArea = sum(
           if_else(substring(PhotoSurface, 2, 2) == "I", TotalArea, 0), na.rm = TRUE),
         TotalTopArea = sum(
           if_else(substring(PhotoSurface, 1, 1) == "T", TotalArea, 0), na.rm = TRUE),
         TotalBottomArea = sum(
           if_else(substring(PhotoSurface, 1, 1) == "B", TotalArea, 0), na.rm = TRUE),
        # Polydora values area
         TotalPolydoraArea = sum(PolydoraArea, na.rm = TRUE),
         PolydoraExternalArea = sum(
           if_else(substring(PhotoSurface, 2, 2) == "E", PolydoraArea, 0), na.rm = TRUE),
         PolydoraInternalArea = sum(
           if_else(substring(PhotoSurface, 2, 2) == "I", PolydoraArea, 0), na.rm = TRUE),
         PolydoraTopArea = sum(
           if_else(substring(PhotoSurface, 1, 1) == "T", PolydoraArea, 0), na.rm = TRUE),
         PolydoraBottomArea = sum(
           if_else(substring(PhotoSurface, 1, 1) == "B", PolydoraArea, 0), na.rm = TRUE),
         .groups = "drop"   # removes grouping after summarise
         ) 

ShellPest2 <- ShellPest1 %>% # getting everything that needs to be scaled
  mutate(TotalWeight_scaled = scale(TotalWeight)[,1],
         ShellHeight_scaled = scale(ShellHeight)[,1],
         Year_scaled = scale(as.numeric(Year))[,1]  # Scaled for better convergence
          ) %>% # getting everything that needs to be factored or numeric
  mutate(Section = factor(Section, levels = c("W", "C", "E")),
         Season = factor(Season, levels = c("Winter", "Spring", "Summer", "Fall")),
         Month = factor(Month, levels = 1:12, labels = month.abb),
         Year = factor(Year),   # Categorical for non-linear trends
         Year_numeric = as.numeric(as.character(Year)),  # For linear trends
         ProcessedBy = factor(ProcessedBy),
         FixedLocation = factor(FixedLocation),
         # Adjust to provide the calculated values for intensity
         PolydoraIntensityWholeOyster = if_else(TotalPolydoraArea == 0, 0, 
                                              round(TotalPolydoraArea / TotalSurfaceArea, 4)),
         PolydoraIntensityExternal = if_else(PolydoraExternalArea == 0, 0, 
                                              round(PolydoraExternalArea / TotalExternalArea, 4)),
         PolydoraIntensityInternal = if_else(PolydoraInternalArea == 0, 0, 
                                              round(PolydoraInternalArea / TotalInternalArea, 4)),
         PolydoraIntensityTop = if_else(PolydoraTopArea == 0, 0, 
                                              round(PolydoraTopArea / TotalTopArea, 4)),
         PolydoraIntensityBottom = if_else(PolydoraBottomArea == 0, 0, 
                                              round(PolydoraBottomArea / TotalBottomArea, 4))
         )

# Create binary infection status (0 = uninfected, 1 = infected)
ShellPest2$Infected <- ifelse(ShellPest2$PolydoraIntensityWholeOyster > 0, 1, 0)

# Remove unnecessary dataframes
rm(hsdbConditionIndex, hsdbShellPest, ShellPest, ShellPest1)

```

<br>~~~ Intensity: Initial GLMM Models ~~~ <br>

``` {r Intensity: Initial GLMM Models}

# Additional optimization options if convergence issues occur:
# control = glmmTMBControl(
#   optimizer = optim,
#   optArgs = list(method = "BFGS"),
#   start = list(theta = log(1)),  # Starting values for dispersion
#   map = list(),  # Fix parameters if needed
#   optCtrl = list(eval.max = 1000, iter.max = 1000)
# )

# Data exploration to see what distribution to use
# ggplot(ShellPest2, aes(x=PolydoraIntensityWholeOyster)) + geom_histogram()
# ggplot(ShellPest2, aes(x=Infected)) + geom_histogram()

##### 1-Section ONLY as a fixed effect
model.S <- glmmTMB(PolydoraIntensityWholeOyster ~ Section,
                   ziformula = ~1,
                   family = beta_family(link = "logit"),
                   data = ShellPest2)

##### 2-TotalWeight ONLY as a fixed effect 
model.TW <- glmmTMB(PolydoraIntensityWholeOyster ~ TotalWeight_scaled,
                   ziformula = ~1,
                   family = beta_family(link = "logit"),
                   data = ShellPest2)

##### 3-ShellHeight ONLY as a fixed effect 
model.SH <- glmmTMB(PolydoraIntensityWholeOyster ~ ShellHeight_scaled,
                   ziformula = ~1,
                   family = beta_family(link = "logit"),
                   data = ShellPest2)

##### 4-Section AND TotalWeight as a fixed effect 
model.S.TW <- glmmTMB(PolydoraIntensityWholeOyster ~ Section + TotalWeight_scaled,
                   ziformula = ~1,
                   family = beta_family(link = "logit"),
                   data = ShellPest2)

##### 5-Section AND ShellHeight as a fixed effect 
model.S.SH <- glmmTMB(PolydoraIntensityWholeOyster ~ Section + ShellHeight_scaled,
                   ziformula = ~1,
                   family = beta_family(link = "logit"),
                   data = ShellPest2)

##### 6-TotalWeight AND ShellHeight as a fixed effect 
model.TW.SH <- glmmTMB(PolydoraIntensityWholeOyster ~ TotalWeight_scaled + ShellHeight_scaled,
                   ziformula = ~1,
                   family = beta_family(link = "logit"),
                   data = ShellPest2)

##### 7-TotalWeight AND ShellHeight AND Section as a fixed effect 
model.TW.SH.S <- glmmTMB(PolydoraIntensityWholeOyster ~ TotalWeight_scaled + ShellHeight_scaled + Section,
                   ziformula = ~1,
                   family = beta_family(link = "logit"),
                   data = ShellPest2)

### Use AICc to compare the 2 models: we'll use the model with the lowest score for making any inferences moving forward
Intensity_models1 <- list(
  '1-SectionOnly' = model.S,
  '2-TotalWeightOnly' = model.TW, 
  '3-ShellHeightOnly' = model.SH, 
  '4-Section+TotalWeight' = model.S.TW, 
  '5-Section+SH' = model.S.SH,
  '6-TotalWeight+SH' = model.TW.SH,
  '7-TotalWeight+SH+Section' = model.TW.SH.S)

aictab(Intensity_models1) #will use one for analysis

# Key findings from model selection
cat("=== Intensity: INITIAL MODEL SELECTION INSIGHTS === 
      1. ShellHeight is highly important (3 >> 1 or 2 alone)
      2. Section is important as it improves both TotalWeight and SH only models (4 & 5 >> 2 & 3)
      3. Including TotalWeight did not provide the best model.")

######### Since Section&TotalWeight has a Massively better AICc, only it will be tested here
best_model1 <- model.S.SH
#########

cat('The following is Residual results, QQ Plot, R2, and Summary for the best fitting model ONLY: 
    5-Section+ShellHeight')

##### Assess goodness-of-fit
simulateResiduals(best_model1, n = 250, refit = FALSE, plot = TRUE)

# Calculate R-squared equivalent for Intensity models. Closer to 1 is better
r2(best_model1)

##### Look at estimates
summary(best_model1)

# Key findings from Initial model 
cat("=== Intensity: INITIAL MODEL INSIGHTS ===
  PolydoraIntensityWholeOyster ~ Section + ShellHeight_scaled
      1. Model fits the data pretty well with good residuals and and R-squared near 1.
      2. Larger oysters have more a higher intensity of polydora. This makes sense and conforms to basic observations.
      3. There are statistically significant differences in the polydora intensity observed in the sections of the Bay.")

# Remove unnecesssary models
rm(model.S, model.S.SH, model.S.TW, model.SH, model.TW, model.TW.SH, model.TW.SH.S)

```

<br>~~~ Intensity: Improving Basic GLMM Models ~~~ <br>
Adding Season and Year to best Initial model

``` {r Intensity: Improving Basic GLMM Models}
# Extended model set: Adding Season and Year systematically

##### 1-Baseline - Best model from above
### Section AND ShellHeight as a fixed effect 
# model.S.SH <- glmmTMB(PolydoraIntensityWholeOyster ~ Section + ShellHeight_scaled,
#                    ziformula = ~1,
#                    family = beta_family(link = "logit"),
#                    data = ShellPest2)

# 2-Add Season (cyclical pattern)
# Section AND TotalWeight AND Season as a fixed effect 
model.S.SH.Ss <- glmmTMB(PolydoraIntensityWholeOyster ~ Section + ShellHeight_scaled + Season,
                   ziformula = ~1,
                   family = beta_family(link = "logit"),
                   data = ShellPest2)

# 3-Add Year as a factor (non-linear temporal trend)
# Section AND TotalWeight AND Year as a fixed effect 
model.S.SH.Y <- glmmTMB(PolydoraIntensityWholeOyster ~ Section + ShellHeight_scaled + Year,
                   ziformula = ~1,
                   family = beta_family(link = "logit"),
                   data = ShellPest2)

# 4-Add Year as a linear trend (more parsimonious)
# Section AND TotalWeight AND Year_scaled as a fixed effect 
model.S.SH.Y_s <- glmmTMB(PolydoraIntensityWholeOyster ~ Section + ShellHeight_scaled + Year_scaled,
                   ziformula = ~1,
                   family = beta_family(link = "logit"),
                   data = ShellPest2)

# 5-Add Season AND Year (factor)
# Section AND TotalWeight AND Season AND Year as a fixed effect 
model.S.SH.Ss.Y <- glmmTMB(PolydoraIntensityWholeOyster ~ Section + ShellHeight_scaled + Season + Year,
                   ziformula = ~1,
                   family = beta_family(link = "logit"),
                   data = ShellPest2)

# 6-Add Season AND Year (linear)
# Section AND TotalWeight AND Season AND Year_Scaled as a fixed effect 
model.S.SH.Ss.Y_s <- glmmTMB(PolydoraIntensityWholeOyster ~ Section + ShellHeight_scaled + Season + Year_scaled,
                   ziformula = ~1,
                   family = beta_family(link = "logit"),
                   data = ShellPest2)

# 7-Add Season-Year interaction (test if seasonal patterns change over years)
# Section AND TotalWeight AND Season-Year_scaled interaction as a fixed effect 
model.S.SH.Ss.X.Y_s <- glmmTMB(PolydoraIntensityWholeOyster ~ Section + ShellHeight_scaled + Season * Year_scaled,
                   ziformula = ~1,
                   family = beta_family(link = "logit"),
                   data = ShellPest2)

### Use AICc to compare the 2 models: we'll use the model with the lowest score for making any inferences moving forward
Intensity_models2 <- list('1-Section+ShellHeight (Base)' = best_model1, 
                 '2-SEC+SH+Season' = model.S.SH.Ss, 
                 '3-SEC+SH+Year' = model.S.SH.Y, 
                 '4-SEC+SH+YearScaled' = model.S.SH.Y_s, 
                 '5-SEC+SH+Season+Year' = model.S.SH.Ss.Y,
                 '6-SEC+SH+Season+YearScaled' = model.S.SH.Ss.Y_s,
                 '7-SEC+SH+Season*YearScaled' = model.S.SH.Ss.X.Y_s
)

aictab(Intensity_models2) #will use one for analysis

# Key findings from model selection
cat("=== Intensity: IMPROVED MODEL SELECTION INSIGHTS ===
      1. Seasonal patterns are not consistent across years (7 >> 5: interaction needed)
      2. Both Season AND Year are highly important (5 >> 2 or 3 alone)
      3. Linear and Non-linear year effects are indistinguishable (5 = 6: factor vs linear year)
      4. Temporal factors explain MASSIVE variation (ΔAICc = 544 vs baseline)")

######### Since 7-SEC+SH+Season*YearScaled has a significantly better AICc, only it will be tested here
best_model2 <- model.S.SH.Ss.X.Y_s
#########

cat('Following is Residual results, QQ Plot, ACF series residuals, R2, and Summary for the best fitting model ONLY:
    7-SEC+SH+Season*YearScaled')

##### Assess goodness-of-fit
simulateResiduals(best_model2, n = 250, refit = FALSE, plot = TRUE)

# Check for temporal autocorrelation since we have time series data
acf(residuals(best_model2, type = "pearson"))
# The residuals exhibit significant autocorrelation, implying that the model has NOT effectively captured the temporal dependencies.

# Calculate R-squared equivalent for Intensity models. Closer to 1 is better
r2(best_model2)

##### Look at estimates
summary(best_model2)

# Key findings from Improved model 
cat("=== Intensity: IMPROVED MODEL INSIGHTS ===
  PolydoraIntensityWholeOyster ~ Section + ShellHeight_scaled + Season * Year_scaled
      1. Model fits the data pretty well with good residuals and and R-squared even closer to 1 than the Initial model.
      2. ACF series residuals series exhibit SIGNIFICANT autocorrelation, implying that the model has NOT captured the temporal dependencies. This is a major issue to resolve before any other insights can be drawn.")

# Remove unnecessary models

rm(model.S.SH.Ss, model.S.SH.Y, model.S.SH.Y_s, model.S.SH.Ss.Y, model.S.SH.Ss.Y_s, model.S.SH.Ss.X.Y_s)
```

<br>~~~ Intensity: Bias Testing ~~~ <br>
Examining the impact of observers and locations, as both have changed over the years


``` {r Intensity: Bias Testing}

# BIAS TESTING MODELS

##### 1-Baseline - Best model from above
### Add Season-Year interaction (test if seasonal patterns change over years)
# # Section AND TotalWeight AND Season-Year_scaled interaction as a fixed effect 
# model.S.SH.Ss.X.Y_s <- glmmTMB(PolydoraIntensityWholeOyster ~ Section + ShellHeight_scaled + Season * Year_scaled,
#                    ziformula = ~1,
#                    family = beta_family(link = "logit"),
#                    data = ShellPest2)

# OBSERVER BIAS TESTING 
# 2-Add ProcessedBy to Intensity model
model.Intensity_observer <- glmmTMB(PolydoraIntensityWholeOyster ~ Section + ShellHeight_scaled + Season * Year_scaled + ProcessedBy,
                            ziformula = ~ 1,
                            family = beta_family(link = "logit"),
                            data = ShellPest2)

# LOCATION BIAS TESTING
# 3-Add FixedLocation as random effect (nested within Section) to Intensity model
model.Intensity_location <- glmmTMB(PolydoraIntensityWholeOyster ~ Section + ShellHeight_scaled + Season * Year_scaled + (1|FixedLocation),
                            ziformula = ~ 1,
                            family = beta_family(link = "logit"),
                            data = ShellPest2)

# COMBINED BIAS TESTING 
# 4-Both observer and location effects on Intensity
model.Intensity_both <- glmmTMB(PolydoraIntensityWholeOyster ~ Section + ShellHeight_scaled + Season * Year_scaled + ProcessedBy + (1|FixedLocation),
                        ziformula = ~ 1,
                        family = beta_family(link = "logit"),
                        data = ShellPest2)

### Use AICc to compare the 2 models: we'll use the model with the lowest score for making any inferences moving forward
# For Intensity models
Intensity_bias_models <- list(
  "Original" = best_model2,  # Your best Improved model
  "Observer_Bias" = model.Intensity_observer,
  "Location_Bias" = model.Intensity_location, 
  "Both_Biases" = model.Intensity_both
)

aictab(Intensity_bias_models)

# Key findings from model selection
cat("=== Intensity: BIAS TESTING MODEL SELECTION INSIGHTS ===
      1. Both Biases are present in Polydora Intensity data (ΔAICc = 928 vs baseline)
      2. Location and Observer have a MASSIVE impact on results")

# simulateResiduals(model.Intensity_observer, n = 150, refit = FALSE, plot = TRUE)
# simulateResiduals(model.Intensity_location, n = 150, refit = FALSE, plot = TRUE)
simulateResiduals(model.Intensity_both, n = 150, refit = FALSE, plot = TRUE)

# Check for temporal autocorrelation since we have time series data
acf(residuals(model.Intensity_both, type = "pearson"))

# r2(model.Intensity_observer)
# r2(model.Intensity_location)
r2(model.Intensity_both)

summary(model.Intensity_both)

# Key findings from Bias testing model 
cat("=== Intensity: BIAS TESTING MODEL INSIGHTS ===
  PolydoraIntensityWholeOyster ~ Section + ShellHeight_scaled + Season * Year_scaled + ProcessedBy + (1|FixedLocation)
      1. Model fits the data pretty well. KS Test fail is expected with a large data set, and given the heavy tails, boundary effects and biases, it's not surprising that KS and outlier tests flag issues.
      2. ACF series residuals series improved, implying that biases were driving the temporal signals, and once controlled for those signals have disappeared. 
      3. A massive drop in R2 (0.921 to 0.038). The high R2 was likely inflated. Once bias is accounted for there isn't much left for the model to explain, thus the explanatory power is very low.
      3. The MASSIVE increase in model performance with both biases raise a lot of doubt about the Improved Model's insights. ")

# Assign predictors of variance to see what has the most amount of influence in the bias model
predictors <- c("Section", "ShellHeight_scaled", "Season * Year", "ProcessedBy")

# Uses Intraclass Correlation Coefficient (ICC) to quantify how much variation in your outcomeis attributable to Random effects
cat("Intraclass Correlation Coefficient (ICC) quantifies how much variation is attributable to Random effects:")
icc(model.Intensity_both)

# uses the importantance_table function to make a clean table detail each factor's importance in the final model
importance_results <- importance_table(model.Intensity_both, predictors) 
kable(importance_results, format = "pipe", digits = 3)

# Key findings from Bias Importance  
cat("=== Intensity: BIAS TESTING IMPORTANCE TABLE & ICC ===
      1. Adjusted ICC indicates that after controliing for other factors, ~2% of the variance is attributable to the random effect of FixedLocation.
      2. Large Positive ΔAIC indicates that ShellHeight and ProcessedBy are extremely important predictors. 
      3. Section's negative ΔAIC indicates it doesn't really improve the model.
      4. ShellHeight, ProcessedBy, and Season*Year have very small p-values (e-157, e-51, and e-37), indicating they are highly significant.
      5. Section is not significant (p = 0.23).
      7. ΔR2 measures how much of the model’s explained variance is uniquely attributable to that predictor. TotalWeight explains ~8% on its own; ProcessedBy adds ~3%; Season*Year adds ~2%; Section contributes virtually nothing.")

cat(" |||||||||||||||||||||||||||| CONCLUSION |||||||||||||||||||||||||||| 
    1. Shell Height is by far the strongest driver of polydora intensity. Larger oysters tend to have more intense polydora.
    2. Given the outsized impact of bias on this data, no other insights can be drawn without significant reprocessing, retraining, and robust QA measures being put in place. After which all these tests need to be run again and more detailed analysis run to determine any significant differences.
    3. This solution represents a large input of staff time (thousands of hours) for a small amount of benefit.
*** 4. Recommendation is to discontinue measuring polydora intensity in ShellPest data. ***")

```

<br>~~~ Prevalence: Data Exploration ~~~ <br>
Basic data exploration of Prevalence data revealed that 4,312 out of 4,395 (98.1%) oysters had some level of polydora infestation. Given the high level of prevalence, no models were fitted to this data, as we are interested more in biologically meaningful patterns and not mere statistical patterns

``` {r Prevalence: Data Exploration}
# Summarizing the mean prevalence temporally (Year and Season)
prevalence_summary1 <- ShellPest2 %>%
  group_by(Year, Season) %>%
  summarise(
    n = n(),
    prevalence = mean(Infected, na.rm = TRUE),
    se = sqrt(prevalence * (1 - prevalence) / n)
  )

# Summarizing the mean prevalence Spatially (Section)
prevalence_summary2 <- ShellPest2 %>%
  group_by(Section) %>%
  summarise(
    n = n(),
    prevalence = mean(Infected, na.rm = TRUE),
    se = sqrt(prevalence * (1 - prevalence) / n)
  )

kable(prevalence_summary1, format = "pipe", digits = 3, label = "Temporal means and standarda errors of polydora prevalence.")

# Key findings from temporal Summary selection
cat("=== Prevalence: TEMPORAL INSIGHTS === 
      1. Most years and seasons have greater than 98% prevalence.
      2. The only 2 seasons with a lower prevalence are in 2016. Given the large impact of observer bias on intensity data, it seems very possible that this is an artifact of bias.")

kable(prevalence_summary2, format = "pipe", digits = 3, label = "Spatial means and standarda errors of polydora prevalence.")

# Key findings from temporal Summary selection
cat("=== Prevalence: SPATIAL INSIGHTS === 
      1. All Sections have greater than 97% prevalence.
      2. There does not appear to be any meaningful biological differences between sections of the bay.")

cat(" |||||||||||||||||||||||||||| CONCLUSION |||||||||||||||||||||||||||| 
    1. There does not appear to be any meaningful biological differences of prevalence on spatial or temporal scales.
    2. A better measure of importance would be measuring the intensity of infestation. However, as noted above there are serious bias issues that would require immense amounts of work to resolve.
*** 3. Recommendation is to discontinue noting the presence or absence of polydora in ShellPest data. ***")

```
