---
title: "oyster_simulation_corrected_202605_MD"
output: html_document
---

### Adapting Colin's code from oyster_simulation_corrected_041626.R 
### Author: Matt Davis

```{r PackageLoad, echo = FALSE, warning = FALSE, message = FALSE}
# Load necessary R packages
library(glmmTMB)
library(tidyverse)
library(DHARMa)
library(cowplot)
library(AICcmodavg)
library(DT)

```

```{r Load Data}
###### Read in data
dater <- read.csv("All_Oysters_2025-05 csv.csv") %>% 
  mutate(FixedLocationID = factor(FixedLocationID), 
         StationName = factor(StationName))

###### Take a quick look at the data
glimpse(dater)

###### Summarize data: count up number of samples, total number of legal oysters, mean, SD, lower 95% CL, upper 95% CL, and CV of the number of legal oysters, and finally the number of 0 observations by FixedLocationID
datSumm <- dater %>% 
  group_by(FixedLocationID) %>% 
  summarize(nSamples = n(), 
            TotalLegal = sum(NumLegal), 
            MeanLegal = round(mean(NumLegal),2), 
            sdLegal = round(sd(NumLegal),2), 
            lwr = round(quantile(NumLegal, 0.025),2), 
            upr = round(quantile(NumLegal, 0.975),2), 
            CVLegal = round(sdLegal/MeanLegal*100,2), 
            prop0 = round(mean(NumLegal==0),2)
            )

###### Look at the summaries
datatable(datSumm)